---
title: "Two way ANOVA"
author: "Pooja Radhakrishnan"
date: "14 April 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
help(crabs)
library(MASS)
names(crabs)
dim(crabs)
attach(crabs)
table(sp,sex)
crab_df <- crabs
```
Problem Statement: Should we consider the variables gender and species to be the deterministic factors of crab morphology?

Here, the categorical variables are: gender and species. There are many response variables namely: Body Depth (BD), Carapace (#the hard shell on top of crab) Length(CL), Carapace Width (CW), Frontal Lobe Size (FL), Rear Width (RW). Here, for performing two-way ANOVA, I pick Body Depth (BD) as my response variable.

But before proceeding to implement the two-way ANOVA, there are a few requisites to be met by the variables:

a) The outcome/response vairable must be continuous.
```{r}
str(crab_df$BD)
```
So, yes, the response variable is continuous.

b) The predictor variables must be categorical.
```{r}
is.factor(crab_df$sp)
is.factor(crab_df$sex)
levels(crab_df$sp)
levels(crab_df$sex)
```
Yes, the predictor variables are categorical.

To check how the variables are closely related to each other, we perform the following to determine the correlation:

Determine correlation from plot:
```{r}
plot(data.frame(BD,CL,CW,FL,RW))
```

Determine correation uing *cor()* function statistically:
```{r}
cor(cbind(BD,CL,CW,FL,RW))
cat("All the values are closer to 1, which means they are closely related and behave similarly.")
```

Checking the values of mean by grouping the categorical and the response variable individually:
```{r}
tapply(BD, sex, mean)
tapply(BD, sp, mean)
```

Checking the value of mean by grouping botht he categorical predictor variables and the response variable:
```{r}
tapply(BD, list(sex,sp), mean)
```

The grand mean of the response variable is:
```{r}
mean(BD)
```

#**Implementation of ANOVA**
Now let us implement ANOVA technique to determine if the cetegorical variables gender adn species have significant effect in determining the response variable Body Depth.

Since there are two categorical variables and one continuous variable, a two-way ANOVA is used.

There are two variants of two-way ANOVA:
* Two-way ANOVA with interaction
* Two-way ANOVA without interaction

#Two way ANOVA with interaction:
We perform this to check if there's a relation between both the categorical variables to determine the response variable.
```{r}
m2i = aov(BD ~ sex*sp) 
summary(m2i)
coef(m2i)
plot(m2i)
```

#Two way ANOVA without interaction:
We perform this variant to check individually about the effect of the predictor variables onthe response variable.
```{r}
m2n = aov(BD ~ sex+sp)
summary(m2n)
coef(m2n)
plot(m2n)
```
From the summaries of both the variant models, both gender and species together, i.e. two-way ANOVA with interaction have significant effect on the response variable. Species alone has a much more significant effect on Body Depth.

The residuals of both the variants of two-way ANOVA follow normal distribution as the points lie close to the reference line.

We can assess the normality by using *shapiro.test()*
```{r}
shapiro.test(crab_df$BD)
```
The p-value is > 0.05. Therefore, normality is assured.

